Family Fortunes: A Quantitative Analysis of the Early Iron Age Cemeteries at Knossos, Crete

### 2019-03-29 analysis/paper/paper.Rmd
writing edits

A first important step in preparing the data for analysis was the estimation of the number of individuals buried in each tomb. The method chosen for this estimation broadly follows that of Cavanagh [-@Cavanagh1996, 659-660], who took the number of pithoi (the most common urn type) and the number of pithoi, amphorae and kraters (which were in some cases used as urns) as middle and upper estimates of the number of interments, with all securely identified interments (in the form of ashes or osteological material) being the lower limit. In the present study, the mean of these three estimates was taken, to provide a single aggregate measure for the purposes of analysis (where no interments were known, this variable was still assigned a value of "1", for all tombs must have contained at least one burial). For analysis at the level of individual finds, however, pithoi alone were taken as the best proxy (Figure \@ref(fig:burialestimate)). For the bulk of the analyses, any tombs lacking identified remains or a single vessel suitable for use as an urn were not considered, thus excluding the most heavily damaged and looted tombs.

@@

library(extrafont)
library(ggplot2)
library(RColorBrewer)
font_import(paths="../paper/font",pattern="GARA")
loadfonts(device="win")

@@

tombs<-read.csv("../data/raw_data/KnossosTombs.csv",header=T)
attach(tombs)

@@

p<-ggplot(tombs)+geom_point(aes(est_burials,all_urns,color="All urns"))+
  geom_point(aes(est_burials,total_burials,color="Known interments"))+
    geom_point(aes(est_burials,pithoi,colour="Pithoi"))+
      geom_line(aes(est_burials,est_burials),alpha=0.4)+
       xlim(0,60)+ylim(0,60)+theme_minimal()+theme(legend.title=element_blank(),
        text=element_text(family="Garamond", size=12))+
        labs(x="Mean Burial Estimate", y="Interment/Vessel Count")+
          scale_color_brewer(palette = "Set2")

@@

cor_burial_proxy<-cor.test(all_urns,est_burials, method = "pearson")

@@

The end

```{r a-table}

@@

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- readr::read_csv(here::here('analysis', 'data', 'raw_data', 'my_csv_file.csv'))
```

>>>>

A first important step in preparing the data for analysis was the estimation of the number of individuals buried in each tomb. The method chosen for this estimation broadly follows that of Cavanagh [-@Cavanagh1996, 659-660], who took the number of pithoi (the most common urn type) and the number of pithoi, amphorae and kraters (which were in some cases used as urns) as middle and upper estimates of the number of interments, with all securely identified interments (in the form of ashes or osteological material) being the lower limit. In the present study, the mean of these three estimates was taken, to provide a single aggregate measure for the purposes of analysis (where no interments were known, this variable was still assigned a value of "1", for all tombs must have contained at least one burial).

@@

library(reshape)
library(data.table)
library(extrafont)
library(ggplot2)
library(RColorBrewer)
library(grid)
library(dplyr)
font_import(paths="../paper/font",pattern="GARA")
loadfonts(device="win")

@@

tombs<-read.csv("../data/raw_data/KnossosTombs.csv",header=T)
tombs<-subset(tombs,tombs$cem_condensed=="NC"|tombs$cem_condensed=="FOR")

@@

p<-ggplot(tombs)+geom_point(aes(est_burials,all_urns,color="All urns"))+
  geom_point(aes(est_burials,total_burials,color="Known interments"))+
    geom_point(aes(est_burials,pithoi,colour="Pithoi"))+
      geom_line(aes(est_burials,est_burials),alpha=0.4)+
       xlim(0,60)+ylim(0,60)+theme_minimal()+
        theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
              labs(x="Mean Burial Estimate", y="Interment/Vessel Count")+
                 scale_color_brewer(palette = "Set2")

@@

cor_burial_proxy<-cor(tombs$all_urns,tombs$est_burials, method = "pearson")

@@

For analysis at the level of individual finds, however, pithoi alone were taken as the best proxy, as they correlated most closely with the mean estimate, producing a pearson correlation coefficient of `r cor_burial_proxy` (Figure \@ref(fig:burialestimate)). For the bulk of the analyses, any tombs lacking identified remains or a single vessel suitable for use as an urn were not considered, thus excluding the most heavily damaged and looted tombs.

```{r aorist_all_cems}

aorist_all_tombs <- colSums(main_tombs[157:206])
aorist_all_tombs<-data.frame(aorist_all_tombs)

colnames(aorist_all_tombs)<-c("aoristic_density")
dates<-seq(from = 610, to = 1100, by = 10)
dates<-rev(dates)
row.names(aorist_all_tombs)<-dates
aorist_all_tombs<-data.frame(aorist_all_tombs)
setDT(aorist_all_tombs, keep.rownames = "decade")
aorist_all_tombs$decade<-as.numeric(aorist_all_tombs$decade)

ggplot(aorist_all_tombs,aes(decade,aoristic_density))+
  geom_col()+xlim(1100,630)+theme_minimal()+
        theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
              labs(x="Date (Years BC)", y="Aoristic Density")+
                 scale_fill_brewer(palette = "Set2")

```{r aorist_uncertainty}

pots<-read.csv("../data/raw_data/KnossosPots.csv",header=T)
pots<-subset(pots,pots$cemetery=="FOR"|pots$cemetery=="TK"|pots$cemetery=="NC")

aorist_pot_values<-pots[24:73]
aorist_pot_matrix<-data.matrix(aorist_pot_values)
aorist_pot_matrix[aorist_pot_matrix == 0] <- NA
aorist_pot_medians<-apply(aorist_pot_matrix,2,median,na.rm=TRUE)

aorist_pot_medians<-data.frame(aorist_pot_medians)
colnames(aorist_pot_medians)<-c("aoristic_median")
row.names(aorist_pot_medians)<-dates
aorist_pot_medians<-data.frame(aorist_pot_medians)
setDT(aorist_pot_medians, keep.rownames = "decade")
aorist_pot_medians$decade<-as.numeric(aorist_pot_medians$decade)

View(aorist_pot_medians)

ggplot(NULL)+geom_col(data=aorist_pot_medians,aes(decade,aoristic_median))+
    xlim(1100,630)+theme_minimal()+
      theme(legend.title=element_blank(),
        text=element_text(family="Garamond", size=12))+
            labs(x="Date (Years BC)", y="Aoristic Median")+
               scale_fill_brewer(palette = "Set2")

ggplot(main_tombs,aes(total_finds))+geom_histogram()+xlim(0,200)

aorist_tomb_values<-tombs[c(1,157:206)]
aorist_tomb_values[aorist_tomb_values == 0] <- NA

aorist_tomb_values<-data.frame(aorist_tomb_values)
colnames(aorist_tomb_values)[2:51]<-dates

aorist_melt<-melt(aorist_tomb_values)
aorist_melt$variable<-as.numeric(as.character(aorist_melt$variable))

aorist_melt$period<-ifelse(aorist_melt$variable<=1100 & aorist_melt$variable>=980,"SM",
		ifelse(aorist_melt$variable<=970 & aorist_melt$variable>=930,"EPG",
		ifelse(aorist_melt$variable<=920 & aorist_melt$variable>=880,"MPG",
		ifelse(aorist_melt$variable<=870 & aorist_melt$variable>=850,"LPG",
		ifelse(aorist_melt$variable<=840 & aorist_melt$variable>=820,"PGB",
	  ifelse(aorist_melt$variable<=810 & aorist_melt$variable>=800,"EG",
	  ifelse(aorist_melt$variable<=790 & aorist_melt$variable>=750,"MG",
	  ifelse(aorist_melt$variable<=740 & aorist_melt$variable>=720,"LG",
	  ifelse(aorist_melt$variable<=710 & aorist_melt$variable>=680,"EO",
	  ifelse(aorist_melt$variable<=670 & aorist_melt$variable>=640,"LO",
	  ifelse(aorist_melt$variable<=630 & aorist_melt$variable>=600,"A",""
		)))))))))))

decade_ranges<-seq(from = 600, to = 1100, by = 10)
decade_ranges<-rev(decade_ranges)
decade_ranges<-as.factor(decade_ranges)
aorist_melt$variable<-as.factor(aorist_melt$variable)
aorist_melt$variable<-factor(aorist_melt$variable, levels=(decade_ranges))
aorist_melt$period<-factor(aorist_melt$period,levels=c("SM", "EPG", "MPG", "LPG", "PGB", "EG", "MG", "LG", "EO", "LO","A"))

ggplot(aorist_melt,aes(variable,value, fill=period))+geom_boxplot(varwidth=TRUE)+theme_bw()+
      theme(legend.title=element_blank(),
        text=element_text(family="Garamond", size=12),
        axis.text.x = element_text(angle = 70, hjust = 1),
        panel.grid.major.x = element_blank())+
            labs(x="Date (Years BC)", y="Aoristic Sums per Tomb")+
               scale_fill_brewer(palette = "Set3")+
                scale_y_continuous(expand = c(0, 0),limits=c(0,20))

```{r aoristknc}
KNC<-subset(tombs,tombs$cem_condensed=="NC")

aorist_KNC <- colSums(KNC[157:206])
aorist_KNC<-data.frame(aorist_KNC)

colnames(aorist_KNC)<-c("aoristic_density")
row.names(aorist_KNC)<-dates
aorist_KNC<-data.frame(aorist_KNC)
setDT(aorist_KNC, keep.rownames = "decade")
aorist_KNC$decade<-as.numeric(aorist_KNC$decade)

FOR<-subset(tombs,tombs$cem_condensed=="FOR")

aorist_FOR <-colSums(FOR[157:206])
aorist_FOR<-data.frame(aorist_FOR)

colnames(aorist_FOR)<-c("aoristic_density")
row.names(aorist_FOR)<-dates
aorist_FOR<-data.frame(aorist_FOR)
setDT(aorist_FOR, keep.rownames = "decade")
aorist_FOR$decade<-as.numeric(aorist_FOR$decade)

ggplot(NULL)+geom_col(data=aorist_all_tombs,aes(decade,aoristic_density,fill="All tombs"))+
  geom_col(data=aorist_KNC,aes(decade,aoristic_density,fill="KNC"))+
  geom_col(data=aorist_FOR,aes(decade,aoristic_density, fill="Fortetsa"))+
      xlim(1100,630)+theme_minimal()+
        theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
              labs(x="Date (Years BC)", y="Aoristic Density")+
                 scale_fill_brewer(palette = "Set2")

```{r aorist_burials}

pithoi<-subset(pots,pots$shape=="Pithos")
all_urns<-subset(pots,pots$shape=="Pithos"|pots$shape=="Amphora"|pots$shape=="Krater")

aorist_pithoi<-colSums(pithoi[24:73])
aorist_pithoi<-data.frame(aorist_pithoi)

colnames(aorist_pithoi)<-c("aoristic_density")
row.names(aorist_pithoi)<-dates
aorist_pithoi<-data.frame(aorist_pithoi)
setDT(aorist_pithoi, keep.rownames = "decade")
aorist_pithoi$decade<-as.numeric(aorist_pithoi$decade)

aorist_all_urns<-colSums(all_urns[24:73])
aorist_all_urns<-data.frame(aorist_all_urns)

colnames(aorist_all_urns)<-c("aoristic_density")
row.names(aorist_all_urns)<-dates
aorist_all_urns<-data.frame(aorist_all_urns)
setDT(aorist_all_urns, keep.rownames = "decade")
aorist_all_urns$decade<-as.numeric(aorist_all_urns$decade)

ggplot(NULL)+geom_col(data=aorist_all_urns,aes(decade,aoristic_density,fill="All urns"))+
  geom_col(data=aorist_pithoi,aes(decade,aoristic_density,fill="Pithoi"))+
      xlim(1100,630)+theme_minimal()+
        theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
              labs(x="Date (Years BC)", y="Aoristic Density")+
                 scale_fill_brewer(palette = "Set2")

```{r aorist_tomb_adjust}

no_tombs<-colSums(main_tombs[157:206]!=0)
no_tombs<-data.frame(no_tombs)
colnames(no_tombs)<-c("no_tombs")
row.names(no_tombs)<-dates
no_tombs<-data.frame(no_tombs)
setDT(no_tombs, keep.rownames = "decade")
no_tombs$decade<-as.numeric(no_tombs$decade)

aorist_all_tombs1<-aorist_all_tombs
aorist_by_tomb<-merge(aorist_all_tombs1,no_tombs)

aorist_by_tomb$aorist_by_tomb<-aorist_by_tomb$aoristic_density/aorist_by_tomb$no_tombs

ggplot(aorist_by_tomb,aes(decade,aorist_by_tomb))+geom_col()+xlim(1100,600)

#Now for the 9th century tombs

tombs_9th<-subset(tombs,tombs$century_built=="9th")
no_tombs_9th<-colSums(tombs_9th[157:206]!=0)

no_tombs_9th<-data.frame(no_tombs_9th)
colnames(no_tombs_9th)<-c("no_tombs")
row.names(no_tombs_9th)<-dates
no_tombs_9th<-data.frame(no_tombs_9th)
setDT(no_tombs_9th, keep.rownames = "decade")
no_tombs_9th$decade<-as.numeric(no_tombs_9th$decade)

aorist_tombs_9th<-colSums(tombs_9th[157:206])
aorist_tombs_9th<-data.frame(aorist_tombs_9th)
colnames(aorist_tombs_9th)<-c("aoristic_density")
row.names(aorist_tombs_9th)<-dates
aorist_tombs_9th<-data.frame(aorist_tombs_9th)
setDT(aorist_tombs_9th, keep.rownames = "decade")
aorist_tombs_9th$decade<-as.numeric(aorist_tombs_9th$decade)

aorist_by_tomb_9th<-merge(aorist_tombs_9th,no_tombs_9th)
aorist_by_tomb_9th$aorist_by_tomb<-aorist_by_tomb_9th$aoristic_density/aorist_by_tomb_9th$no_tombs

ggplot()+geom_col(data=aorist_by_tomb_9th,aes(decade,aorist_by_tomb,fill="9th Century Tombs"))+
  geom_col(data=aorist_by_tomb,aes(decade,aorist_by_tomb, fill="All Tombs"),alpha=0.7)+
    xlim(1100,600)+theme_minimal()+
        theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
              labs(x="Date (Years BC)", y="Aoristic Density/Number of Tombs Represented")+
                 scale_fill_brewer(palette = "Set2")

names(main_tombs)
tombs_in_use<-colSums(main_tombs[5:14]!="")
tombs_in_use<-data.frame(tombs_in_use)
colnames(tombs_in_use)<-c("no_tombs")
tombs_in_use<-data.frame(tombs_in_use)
setDT(tombs_in_use, keep.rownames = "period")
tombs_in_use$period<-factor(tombs_in_use$period,levels=c("SM", "EPG", "MPG", "LPG", "PGB", "EG", "MG", "LG", "EO", "LO"))

ggplot(no_tombs,aes(decade,no_tombs, fill="Tombs in Use"))+geom_col()+theme_classic()+
     theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
            xlim(1100,600)+
              labs(x="Date (Years BC)", y="Number of Tombs in Use")+
                 scale_fill_brewer(palette = "Set2")

ggplot(tombs_in_use,aes(period,no_tombs, fill="Tombs in Use"))+geom_col()+theme_classic()+
        theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
              labs(x="Period", y="Number of Tombs in Use")+
                 scale_fill_brewer(palette = "Set2")

ggplot(tombs,aes(start_period, fill="New Tombs"))+geom_bar()+theme_classic()+
        theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
              labs(x="Period", y="Number of New Tombs")+
                 scale_fill_brewer(palette = "Set2")

tombs$start_period<-factor(tombs$start_period,levels=c("SM", "EPG", "MPG","PG", "LPG", "PGB", "EG", "MG","G", "LG", "EO","O", "LO"))

```{r rates of burials}

tombs_ranked<-main_tombs[with(main_tombs, order(-main_tombs$total_pottery)), ]
tombs_top_12<-head(tombs_ranked,12)

no_tombs_top<-colSums(tombs_top_12[157:206]!=0)

no_tombs_top<-data.frame(no_tombs_top)
colnames(no_tombs_top)<-c("no_tombs")
row.names(no_tombs_top)<-dates
no_tombs_top<-data.frame(no_tombs_top)
setDT(no_tombs_top, keep.rownames = "decade")
no_tombs_top$decade<-as.numeric(no_tombs_top$decade)

aorist_tombs_top<-colSums(tombs_top_12[157:206])
aorist_tombs_top<-data.frame(aorist_tombs_top)
colnames(aorist_tombs_top)<-c("aoristic_density")
row.names(aorist_tombs_top)<-dates
aorist_tombs_top<-data.frame(aorist_tombs_top)
setDT(aorist_tombs_top, keep.rownames = "decade")
aorist_tombs_top$decade<-as.numeric(aorist_tombs_top$decade)

aorist_by_tomb_top<-merge(aorist_tombs_top,no_tombs_top)
aorist_by_tomb_top$aorist_by_tomb<-aorist_by_tomb_top$aoristic_density/aorist_by_tomb_top$no_tombs

ggplot()+geom_col(data=aorist_by_tomb_top,aes(decade,aorist_by_tomb,fill="12 Largest Tombs"))+
  geom_col(data=aorist_by_tomb,aes(decade,aorist_by_tomb, fill="All Tombs"),alpha=0.7)+
    xlim(1100,600)+theme_minimal()+
        theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
              labs(x="Date (Years BC)", y="Aoristic Density/Number of Tombs Represented")+
                 scale_fill_brewer(palette = "Set2")

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- readr::read_csv(here::here('analysis', 'data', 'raw_data', 'my_csv_file.csv'))

@@
